### 2024-05-15 2024-5-13/sections/cefa.tex
commit the paper extended by setta2023

The semantics of CEFA is defined as follows. Let $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$ be a CEFA. 
A \emph{run} of $\aut$ on a string $w = a_1 \cdots a_n$ is a sequence $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ such that $q_0 \in I$ and $q_{i-1} \xrightarrow[\myvec{v_i}]{a_i} q_i$ for each $i \in [n]$. A run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ is \emph{accepting} if $q_n \in F$ and $\alpha(\myvec{v'}/R)$ is true, where $\myvec{v'} = \sum \limits_{j \in [n]} \myvec{v_j}$. The vector $\myvec{v'}= \sum \limits_{j \in [n]} \myvec{v_j}$ is called the \emph{cost} of an accepting run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$. Note that the values of all registers are initiated to zero and updated to $\sum \limits_{j \in [n]} \myvec{v_j}$ after all the transitions in the run are executed. We use $\myvec{v'} \in \aut(w)$ to denote the fact that there is an accepting run of $\aut$ on $w$ whose cost is $\myvec{v'}$.  
We define the semantics of a CEFA $\aut$, denoted by $\Lang(\aut)$, as $\{(w; \myvec{v'}) \mid  \myvec{v'} \in \aut(w)\}$.  In particular, if $I \cap F \neq \emptyset$ and $\alpha[\myvec{0}/R]$ is true, then $(\varepsilon; \myvec{0}) \in \Lang(\aut)$. Moreover, we define the \emph{output} of a CEFA $\aut$, denoted by $\cefaout(\aut)$, as $\{\myvec{v'} \mid  \exists w.\ \myvec{v'} \in \aut(w)\}$.

>>>>

The semantics of CEFA is defined as follows. Let $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$ be a CEFA. 
A \emph{run} of $\aut$ on a string $w = a_1 \cdots a_n$ is a sequence $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ such that $q_0 \in I$ and $q_{i-1} \xrightarrow[\myvec{v_i}]{a_i} q_i$ for each $i \in [n]$. A run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ is \emph{accepting} if $q_n \in F$ and $\alpha(\myvec{v'}/R)$ is true, where $\myvec{v'} = \sum \limits_{j \in [n]} \myvec{v_j}$. The string $w$ is $accepted$ by the CEFA if and only if there is an accepting run of $w$. The vector $\myvec{v'}= \sum \limits_{j \in [n]} \myvec{v_j}$ is called the \emph{cost} of an accepting run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$.Note that the values of all registers are initiated to zero and updated to $\sum \limits_{j \in [n]} \myvec{v_j}$ after all the transitions in the run are executed. We use $\myvec{v'} \in \aut(w)$ to denote the fact that there is an accepting run of $\aut$ on $w$ whose cost is $\myvec{v'}$.  
We define the semantics of a CEFA $\aut$, denoted by $\Lang(\aut)$, as $\{(w; \myvec{v'}) \mid  \myvec{v'} \in \aut(w)\}$.  In particular, if $I \cap F \neq \emptyset$ and $\alpha[\myvec{0}/R]$ is true, then $(\varepsilon; \myvec{0}) \in \Lang(\aut)$. Moreover, we define the \emph{output} of a CEFA $\aut$, denoted by $\cefaout(\aut)$, as $\{\myvec{v'} \mid  \exists w.\ \myvec{v'} \in \aut(w)\}$.

### 2024-05-15 2024-5-13/sections/introduction.tex
commit the paper extended by setta2023

Counting (aka repetition) is a convenient feature in regexes that count the number of matchings of sub-patterns. For instance, $a^{\{2, 4\}}$ specifies that $a$ occurs at least twice and at most four times, and $a^{\{2, \infty\}}$ specifies that $a$ occurs at least twice. 
Note that the Kleene star and the Kleene plus operator are special cases of counting. For instance, $a^*$ is equivalent to $a^{\{0,\infty\}}$ and $a^+$ are equivalent to $a^{\{1,\infty\}}$.
Counting is a frequently used feature of regexes. According to the statistics from \cite{CS16}, Kleene star/plus occur in more than 70\% of 1,204 Python projects, while other forms of counting occur in more than 20\% of them. Therefore, an efficient analysis of string manipulating programs requires efficient solvers for string constraints containing regexes with counting\footnote{In the rest of this paper, for clarity, we use counting to denote expressions of the form $e^{\{m, n\}}$ and  $e^{\{m, \infty\}}$, but not $e^*$ or $e^+$.} and string-length function at least.

@@

\medskip
\noindent
\emph{Related work.} 
We discuss more related work on regexes with counting, string-length function, and automata with registers/counters.  
%
Determinism of regexes with counting was investigated in~\cite{GGM12,CL15}. 
%
Real-world regexes in programming languages include features beyond classical regexes, e.g., the greedy/lazy Kleene star, capturing groups, and back references. Real-world regexes have been addressed in symbolic execution of Javascript programs~\cite{LMK19} and string constraint solving~\cite{CF+22}. 
Nevertheless, the counting operators are still unfolded explicitly in~\cite{CF+22}. 
%
The Trau tool focuses on string constraints involving flat regular languages and string-length function and solves them by computing LIA formulas that define the Parikh images of flat regular languages~\cite{z3trau}. The Slent tool solves the string constraints involving string-length function by encoding them into so-called length-encoded automata, then utilizing symbolic model checkers to solve their nonemptiness problem~\cite{WC+18}. However, neither Trau nor Slent supports counting operators explictly, in other words, counting operators in regexes should be unfolded before solved by them.
%
Cost registers in CEFAs are different from registers in (symbolic) register automata~\cite{ra,sra}:  In register automata, registers are used to store input values and  can only be compared for equality/inequality, while in CEFAs, cost registers are used to store integer-valued costs and can be updated by adding/subtracting integer constants and constrained by the accepting conditions which are LIA formulas.
%
Therefore, cost registers in CEFAs are more like counters in counter automata/machines~\cite{Minsky67}, that is, CEFAs can be obtained from counter machines by removing  transition guards and adding accepting conditions. 
%
%For the automata model, the definition of registers of CEFAs resembles the counters of the counter automata class~\cite{Minsky67}, but CEFAs contain no guards in each transition. 
%
Counting-set automata were proposed in~\cite{redos_lenka,HS+23} to quickly match a subclass of regexes with counting. Moreover, a variant of nondeterministic counter automata, called bit vector automata, was proposed recently in \cite{GKM23} to enable fast matching of a more expressive class of regexes with counting.   Nevertheless, the nonemptiness problem of these automata was not considered, and it is unclear whether these automata models can be used for solving string constraints with regex-counting and string-length.

>>>>

Counting (aka repetition) is a convenient feature in regexes that counts the number of matchings of sub-patterns. For instance, $a^{\{2, 4\}}$ specifies that $a$ occurs at least twice and at most four times, and $a^{\{2, \infty\}}$ specifies that $a$ occurs at least twice. 
Note that the Kleene star and the Kleene plus operator are special cases of counting. For instance, $a^*$ is equivalent to $a^{\{0,\infty\}}$ and $a^+$ is equivalent to $a^{\{1,\infty\}}$.
Counting is a frequently used feature of regexes. According to the statistics from \cite{CS16}, Kleene star/plus occur in more than 70\% of 1,204 Python projects, while other forms of counting occur in more than 20\% of them. Therefore, an efficient analysis of string manipulating programs requires efficient solvers for string constraints containing regexes with counting\footnote{In the rest of this paper, for clarity, we use counting to denote expressions of the form $e^{\{m, n\}}$ and  $e^{\{m, \infty\}}$, but not $e^*$ or $e^+$.} and string-length function at least.

@@

\medskip
\noindent
\emph{Related work.} 
We discuss more related work on regexes with counting, string-length function, and automata with registers/counters.  
%
Determinism of regexes with counting was investigated in~\cite{GGM12,CL15}. 
%
Real-world regexes in programming languages include features beyond classical regexes, e.g., the greedy/lazy Kleene star, capturing groups, and back references. Real-world regexes have been addressed in symbolic execution of Javascript programs~\cite{LMK19} and string constraint solving~\cite{CF+22}. 
Nevertheless, the counting operators are still unfolded explicitly in~\cite{CF+22}. 
%
The Trau tool focuses on string constraints involving flat regular languages and string-length function and solves them by computing LIA formulas that define the Parikh images of flat regular languages~\cite{z3trau}. The Slent tool solves the string constraints involving string-length function by encoding them into so-called length-encoded automata, then utilizing symbolic model checkers to solve their nonemptiness problem~\cite{WC+18}. However, neither Trau nor Slent supports counting operators explicitly, in other words, counting operators in regexes should be unfolded before solved by them.
%
Cost registers in CEFAs are different from registers in (symbolic) register automata~\cite{ra,sra}:  In register automata, registers are used to store input values and  can only be compared for equality/inequality, while in CEFAs, cost registers are used to store integer-valued costs and can be updated by adding/subtracting integer constants and constrained by the accepting conditions which are LIA formulas.
%
Therefore, cost registers in CEFAs are more like counters in counter automata/machines~\cite{Minsky67}, that is, CEFAs can be obtained from counter machines by removing  transition guards and adding accepting conditions. 
%
%For the automata model, the definition of registers of CEFAs resembles the counters of the counter automata class~\cite{Minsky67}, but CEFAs contain no guards in each transition. 
%
Counting-set automata were proposed in~\cite{redos_lenka,HS+23} to quickly match a subclass of regexes with counting. Moreover, a variant of nondeterministic counter automata, called bit vector automata, was proposed recently in \cite{GKM23} to enable fast matching of a more expressive class of regexes with counting.   Nevertheless, the nonemptiness problem of these automata was not considered, and it is unclear whether these automata models can be used for solving string constraints with regex-counting and string-length.
